### 2022-06-16 CORL/example.tex
minor things

%TODO: Revise method section
%TODO: Create formulation section
%TODO: Create results section

%TODO:
%	Formulation Section
%	Figures
%		Main diagram - break it up into pieces with just text
%		UVL three-axis diagram
%		Replace C with (pi,s)
%	Method:
%		Unprojection Consistency Loss
%		Texture Reality Loss

%TODO:
%	Formulation Section
%	Figures: Main diagram - break it up into pieces with just text
%	Figures: UVL three-axis diagram
%	Method:	Unprojection Consistency Loss
%	Method:	Texture Reality Loss

%ALL CURRENTLY USED TERMS:

%RULES:
%    - Lower case: Belongs to an upper case thing

%     L - num labels
%     N - batch size
%     T - translator
%     A, B - domains
%     Tba
%     Tab
%     tau - texture       !!!!! MUST CHANGE
%			TO DO: replace it with psi ψ, the triton symbol!
%     pi - projection
%     p - photo
%     s - scene
%     l_TR
%     l_UC
%     l_ADV
%     l_REC
%     l_CYC
%     l_T
%     i - texture index
%     f_i - fourier frequency
%     M - MLP for texture network
%     u,v,l,r,g,b - channel values
%     u - unprojections  !!!!!! MUST CHANGE
%          make it capitalized. Unprojection should be denoted by U - it's too intuitive not to.
%     sigma - standard deviation
%     mu - mean
%     E_A, E_B
%     D_A, D_B
%     G_A, G_B

%     similarity(x,y) = MSSSIM(L2(x,y)

% FIXING NOTATION WILL BE HARD :(
% GO OVER WITH XIANG AND JINGHUAN TOMORROW
%     In this paper I use 1-based indexing to make notation simpler.
%
%     s, pi in N,3,H,W - height width
%              i,c,x,y
%     All values between 0 and 1
%
%     I: {1/L, 2/L, ... L/L} -> {1, 2, ... L}  ;   I(l) = l*L
%          I is a discrete mapping from continous label values l to discrete texture indices.
%          It can be nonlinear, as long as its consistent.
%     
%     ψ_i: R2 -> R3   for   i \in {1,2 ... L}
%
%     pi_{n,c,x,y} = ψ_  {I(s_{n,1,x,y})}  (s_{n,2,x,y},s_{n,3,x,y})
%
%     EXPLAINING L_UC IS WAY TO BURDENSOME. We have to talk about weighted standard deviations of unprojections that have a separate unprojection resolution (128x128) etc.
%	  l_UC = total mean across all dims of sigma
%     sigma \in R^{L x C x 128 x 128} is the standard deviation of u in the channel dimension, weighted...
%     !!! I shouldn't have to do this; the other paper didn't!!!
%     u.sum_{N,C,u,v}
%

\newcommand{\msssim}{\text{msssim}}

\newcommand{\Tab}{T_{A\rightarrow B}}
\newcommand{\Tba}{T_{B\rightarrow A}}
\newcommand{\pih}{\hat{\pi}}
\newcommand{\sh}{\hat{s}}
\newcommand{\ph}{\hat{p}}
\newcommand{\pis}{\pi,s} % pi and s pairs from domain A
\newcommand{\pish}{\pih,\sh} % pi and s pairs from domain A
\newcommand{\ppis}{(\pis)} %Parenthesized pis
\newcommand{\ppish}{(\pish)} %Parenthesized pis

\newcommand{\toN}{^{1\dots N}}
% \newcommand{\toNL}{_{1\dots B,\text{ } 1\dots L}}
\newcommand{\toNL}{^{1\dots N, 1\dots L}}
\newcommand{\toL}{^{1\dots L}}

\newcommand{\lUC}{\ell_{UC}}
\newcommand{\lTR}{\ell_{TR}}
\newcommand{\eqand}{\ \ \ \text{ and }\ \ \ }
\newcommand{\eqwhere}{\ \ \ \text{ where }\ \ \ }
\newcommand{\real}{\mathbb{R}}

% \newcommand{\similar}{\text{similarity}}
\newcommand{\similar}{\Omega}
\newcommand{\unprojection}{\omega}
\newcommand{\expectation}{\mathop{\mathbb{E}}}
\newcommand{\lsum}{\sum\limits}

%%%%%%%%%%%%BIG TODO::::
%%%%%% MAKE BATCH SIZE N INSTEAD OF B - B is a domain!

\documentclass{article}

\usepackage{corl_2022} % Use this for the initial submission.
% \usepackage[final]{corl_2022} % Uncomment for the camera-ready ``final'' version.
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{array, booktabs, ragged2e}

%\usepackage[preprint]{corl_2022} % Uncomment for pre-prints (e.g., arxiv); This is like ``final'', but will remove the CORL footnote.

% \title{TRITON: Globally Consistent Sim-to-Real Transfer
\title{Neural Neural Textures Make Better Simulators

``I Can't Believe It's Not Real!''%Definitely gonna delete this lol
 %via Neural Textures
 }

\author{%
	Ryan Burgert \\
	Department of Computer Science\\
	Stony Brook University\\
	Stony Brook, NY 11794 \\
	\texttt{rburgert@cs.stonybrook.edu} \\
	% \And
	% Michael Ryoo \\
	% Department of Computer Science\\
	% Stony Brook University\\
	% Stony Brook, NY 11794 \\
	% \texttt{mryoo@cs.stonybrook.edu} \\
	\And
	Emma Burgert \\
	Ryan Burgert's Dog\\
}

\begin{document}
\maketitle

%===============================================================================

\begin{abstract}
	% Abstracts should be a single paragraph, between 4--6 sentences long, ideally. Gross violations will trigger corrections at the camera-ready phase.
	% Because of the 6 sentence limit, I've commented out some of the sentences...
	Unpaired image translation algorithms can be used for sim-to-real tasks, but most fail to generate temporally consistent results.
	% Because of this, the details or even identity of a particular object might change as it moves across a scene.
	We present a new approach that combines differential rendering with image translation to achieve temporal consistency over indefinite timescales.
%
	%TURNING TWO SENTENCES INTO ONE:
	% We call this algorithm TRITON (Texture Recovering Image Translation Network): an unsupervised, end-to-end, stateless sim-to-real algorithm.
	% Unlike previous techniques, TRITON leverages the underlying 3d geometry of input scenes by generating realistic-looking learnable neural textures.
	We call this algorithm TRITON (Texture Recovering Image Translation Network): an unsupervised, end-to-end, stateless sim-to-real algorithm that 
	leverages the underlying 3d geometry of input scenes by generating realistic-looking learnable neural textures.
%
	% These textures are projected onto the surfaces of input scenes, which are then fed through an image translation network to obtain the final outputs.
	By settling on a particular albedo for the objects in a scene, we ensure consistency between frames statelessly.
	Unlike previous algorithms, TRITON is not limited to camera movements - it can handle the movement of objects as well, making it useful for downstream tasks such as robotic manipulation.
	Our experiments show that in addition to achieving higher temporal consistency, the translations are closer to ground truth photographs than previous techniques.
\end{abstract}
\keywords{sim to real, image translation, differentiable rendering}

%===============================================================================
\section{Introduction}
Problem: 
Why is this problem important?
No view consistency.

Introducing our intuition of what improvements can solve this problem.

Our exact method - introducing our components and what we're doing

Conclusion: This works well, and our contributions are x

% I introduce you to TRITON. The rest of the introduction text goes here; I've commented out my old incomplete introduction.

	% TODO: Note that the simple input of UVL, or an RGB neural texture, means any simulator should be plug-and-play with this algorithm.

	% Recently, several sim-to-real algorithms have come out that are based on image-to-image translation. [Citations]
	% These are used for training robots using pixel-based reinforcement learning.
	% They demonstrated that better image translation algorithms yielded notable improvements in robotic benchmarks.
	% At the same time, there is a lot of research on differentiable rendering,
	% a topic which produced papers such as NERF which attempt to capture 3d representations of the real world.
	% However, most of these approaches are either supervised or involve only camera movements [citations].
	% One paper that came out recently combined both image translation and differentiable rendering in an unsupervised manner to create view-consistent models of human organs [citation].
	% This paper too, though, was limited to only camera movements.

	% This algorithm unlocks a new capability: learning textures to match photos for given 3d models without supervision.

	% The motivation for this paper is to create a temporally consistent sim-to-real image translation algorithm that will look the same from multiple camera angles simultaneously, keeps all semantic details 
	% [[I'll finish this later... let's work on the meat of the paper first...]]

%===============================================================================
\section{Formulation}
\label{sec:data}
	\begin{figure}[H]
	\vspace{-10pt}
		\begin{center}
			\includegraphics[width=400pt]{../images/first_diagram.pdf}
		\end{center}
		\vspace{-5pt}
		\caption{
			This is an overview of what TRITON accomplishes.
			TRITON learns the textures of 3d objects to help translate simulated images into photographs. It does this with unpaired data. Each row is a different dataset.
			}
		\label{fig:first_diagram}
	\end{figure}
	% TRITON stands for Texture Recovering Image Translation Network. It has two goals: to translate images, and to recover textures.

	TRITON's goal is to turn 3d simulated images into realistic fake photographs (by training it without any matching pairs), while maintaining high surface consistency.
	It does this by simultaneously learning both an image translator and a set of realistic textures.
	These translations can be useful for downstream tasks, especially robotic policy learning from camera data, enabling sim2real.

	In contrast to previous works involving `neural textures' \textbf{(mryoo: cite them here)}, TRITON can handle more than just camera movements:
	the positions of objects in the training data can be moved around or deformed as well between data samples.
	For this reason, we chose the term ``surface consistency'' instead of ``view consistency'', which was the term used in [] \textbf{(mryoo: cite)}.
	With high surface consistency, surfaces of translated objects will look the same even when moved around or viewed from different camera angles.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=400pt]{../images/uvl_explanation_minimal.pdf}
		\end{center}
		\caption{
			A scene is obtained by rendering a 3d state into an image, where the first two color channels represent $u,v$ coordinates and the third channel $l$ represents object labels.
			% This diagram shows the purpose of a UVL image.
			% A UVL scene is an image that has three channels: RGB.
			% R and G are used for U and V, and B is used for L.
			% L stands for Label, and determines which part of the synthetic image will get which texture.
			% The U and V coordinates correspond to positions on each texture,
			% telling it how to project the textures onto the 3d objects represented in the UVL scene.
			% The darkest shade of blue corresponds to the 1st texture, and the next darkest shade of blue corresponds to the 2nd texture etc.
			}
		\label{fig:uvl_explanation}
	\end{figure}

	There are two components to every dataset: A set of simulated 3d scenes, and a set of photographs.
	Datasets usually have many more scenes than photographs, since scenes can be generated automatically.

	% The photographs, having $(r,g,b)$ values between 0 and 1,
	% 	are denoted as $p \in [0,1]^{N \times 3 \times H \times W}$ where $N$ is the batch size,
	% 	3 refers to the three $(r,g,b)$ channels, 
	% 	and $H,W$ are the height and width respectively.

	% A photograph in a set of photographs, having $(r,g,b)$ values between 0 and 1,
	% 	is denoted as ${p^n \in [0,1]^{ 3 \times H \times W}}$ where $n$ is that photograph's index in that set,
	% 	3 refers to the three $(r,g,b)$ channels, 
	% 	and $H,W$ are the height and width respectively.

	A photograph is an image tensor, having $(r,g,b)$ values between 0 and 1,
		is denoted as ${p \in [0,1]^{ 3 \times H \times W}}$ where 
		3 refers to the three $(r,g,b)$ channels, 
		and $H,W$ are the height and width respectively.

	In this work, a scene refers to a rendered 3d simulation state - which includes the position, rotation, and deformation of every object (including the camera).
	% Using some 3d rendering software, be it Blender or Gazebo or some other 3d creates an image from these 
	We represent each 3d state with a simple $(r,g,b)$ image, in a simple enough format that any decent simulator should able to provide.
	% These have the same datatype and dimensions as photographs and are denoted as 
	% $s \in [0,1]^{N \times 3 \times H \times W}$.
	% A scene has the same dimensions as a photograph and is denoted as 
	% $s^n \in [0,1]^{3 \times H \times W}$.
	A scene has the same dimensions as a photograph and is denoted as 
	$s \in [0,1]^{3 \times H \times W}$.
	However, unlike $p$, $s$'s three $(r,g,b)$ channels encode a special semantic meaning: $(u,v,l)$ as depicted in figure \ref{fig:uvl_explanation}.
	The channels $u$ and $v$ refer to a texture coordinate, whereas $l$ refers to a label value that identifies every object in a scene.
	In a given dataset, each object gets a different label. Likewise, each object is assigned a different texture.
	We assume that every dataset has $L$ different label values, $L$ different objects, and that we must learn $L$ different neural textures.

TODO: Say I scales and discretizes l, don't say more than that. 
% 	In each dataset, we must also define a function $I(l)$ that maps label values to texture indices. A simple way to do this is with the following definition:
% 	\begin{equation}
% 		I: \left\{\frac{1}{L},\frac{2}{L},\dots \frac{L}{L}\right\} \rightarrow \left\{1,2,\dots L\right\}  \eqand  I(l)=l\cdot L
% 	\end{equation}
	However, this mapping can be changed arbitrarily if its more convenient to do so for a given dataset.
	Also please note that this paper uses 1-based indexing for notational simplicity.

	% TODO: Describe what UVL is ; 
	% Formulate the setting: What our inputs are and the targets are and what we learn (a neural texture function) and how we can take advantage of it for robotic learning.
	% Don't go into details; just say 'its a black box function'

	% TODO: Add UV to the fourier feature diagram.

	% TODO: Create subfigures with subcaptions. 
	% 2a: Nothing but UVL=UV+L
	% 2b. 3d uvl axis to texture pack (drew on whiteboard)
	% 2c. Skip over the translation network and show a pretty robot arm picture

	% This is the formulatoin section; not data section. Dont talk about how many photos we use etc.

	% JUST IN CASE THINGS ARE ERASED:
	% We learn neural texture corresponding to each object, where each texture is represented as a function mapping... 
	% The idea is to learn object textures as a set of continuous functions from a set of (unlabeled) real images, and use them to generate realistic images.
	% Given a 3D structure of the scene represented with a UVL image, the learned neural textures will translate them into a realistic image.
	% When used for the robot learning, this enables learnable, realistic sim2real of robot training images.

	% IN THE INTRODUCTION: How we're using a neural neural network network texture. Also in related works.

	% TODO: Replace neural texture in figure 3 with axis, and treat it as a function that takes UV in with an arrow pointing to it.

	% BREAK IT DOWN APPROACH:
	% Don't have any pictures. Have breakdown where we add new things in multiple figures.
	% First figure: Just image translation. 
	% Seconf:
	% Third:
	% Etc...

	% In this paper, we'll mostly focus on a hand-collected dataset that involves three alphabet blocks that are moved across a table.
	% Alphabet blocks were chosen because they all have the same geometry (a cube), but have very distinct semantic textures that will make temporal inconsistency obvious to the human eye.
	% This dataset is analogous to data that would be collected for a third-person robotic grasping task.
	% However, other objects (such as cans of soda, apples, cloves of garlic, etc) are also for generating other results in this paper. ((TODO: Detail the five object tests and five cube tests and other stuff like shiny cans etc)).

	% There are two components to this dataset: Simulated data in the form of UVL scenes, and photographs.
	% The camera in the simulation is positioned in approximately the same place as the real camera, with approximately the same field of view.
	% In both domains of this dataset, the camera never moves. Instead, the objects are randomly positioned and rotated on the table.
	% In both the simulation and in real life, these objects are randomly translated along the x and y-axis of the table, as well as rotated on the z-axis.

	% There are 200 photographs in this dataset, and they are recorded as RGB jpg images.
	% There are 3000 synthetic images in this dataset, in the form of UVL images.
	% These UVL images are also encoded as RGB images, but provide information about the geometry of a scene.
	% UVL images are composed of both a UV map and a label map and are summarized in figure \ref{fig:uvl_explanation}.
	% To avoid positional rounding errors, UVL images are encoded as EXR files: an image format that supports floating-point values.

	% The pixels in a UVL image are visualizable in RGB space because, like RGB images, each pixel is a vector with 3 values. The first two channels, U and V, are floating point values between 0 and 1. The L channel is also a floating point value between 0 and 1, it has a second meaning. There are only a finite number of label values (corresponding to the number of learnable textures). The rank of a particular L value is used to index these textures.

	% In the three-cube dataset we have 4 textures, and UVL scenes contain label values $l \in [\frac{0}{3}, \frac{1}{3},\frac{2}{3}, \frac{3}{3}]$. A pixel in a UVL scene $l=\frac{2}{3}$ will get indexed to the third texture, because $\frac{2}{3}$ is the third smallest label value.

	% Domain A and B
	% p hat
	% s hat

% %===============================================================================
\section{Method} \label{sec:method}
	\begin{figure}[thbp]
	    \vspace{-20pt}
		\begin{center}
			% \includegraphics[width=400pt]{../images/main_diagram_condensed.pdf}
			% \includegraphics[width=400pt]{../images/main_diagram_condensed_with_domains.pdf}
			\includegraphics[width=\textwidth]{../images/main_diagram_beautiful.pdf}
		\end{center}
		\vspace{-10pt}
		\caption{
			% We use differentiable rendering paired with unpaired image translation to consistently= transform synthetic scenes into realistic images.
			% To ensure all objects look consistent in different positions and views, we learn the albedo of each object's surface by using a learnable latent texture.
			% This texture is parametrized as an image generated by an MLP which is fed Fourier features, as shown in figure \ref{fig:learnable_textures}.
			% $N$ is the batch size, and $L$ is the number of different labels (which is also the number of learnable textures).
			%
			%JUNE 12:
			%Don't say 4 main components; say 2 components and 2 losses
			%TODO: Try to use right-angled arrows
			%TODO: Use tau as a function
			%Copy paste the figure multiple times, deleting things
			%Mention "Neural Neural Texture": Its a function mapping UV to RGB
			%Formulation doesnt mentino texture; just uvl inputs and final outputs
			%Always use function notation tauuv instead of just tau (in figure as well) 
			%The arrow should be running through tau. It's not a variable its a function...
			%Rename 'B' to "N" for batch size
			%Figure 8: Align consistent with inconsistent and use all 5 green ones
			%
			This figure is an overview TRITON.
			Inputs are on the left, and outputs are on the right.
			The dashed arrows indicate losses.
			Although $\pi\toN$ and $\ph\toN$ look similar, they are not identical - $\lTR$ encourages them to look as similar as possible.
			% $\pi$ is simply projected textures, whereas $\hat{p}$ has been run through an image translator, allowing it to have shadows and specularities.
			% TODO: Come up with a tour of this diagram in this caption. Also maybe make it larger?
			% TODO: Use variables instead of 3,4
			% TODO: Make the text larger and easier to read
			% TODO: Simplify the diagram?
		}
		\label{fig:main_diagram}
	\end{figure}

		In this section, we describe how TRITON works to translate images from domain A (simulation) to domain B (photos) while maintaining surface consistency. 
		TRITON has four main components: a neural texture, an image translator, and two surface consistency losses. [[One or two sentences summarizing the flow - nothing more than that]].

		%JUNE 12:
		%TODO: MIgrate all four of these teasers into their respective subsections to save space.

	\subsection{Neural Neural Texture Projection}
		\begin{figure}[thbp]
		    \vspace{-15pt}
			\begin{center}
				\includegraphics[width=300pt]{../images/learnable_textures.pdf}
			\end{center}
			\vspace{-3pt}
			\caption{
				This figure illustrates how the our neural neural textures are calculated in figure \ref{fig:main_diagram}. They are fully differentiable, and represented continuously along the texture's U and V axes using Fourier feature networks.
			}
			\vspace{-10pt}
			% For each texture, there is a separate Fourier feature network (Fourier features fed into a pixel-wise MLP) that will take in Fourier features as inputs, and output an RGB image using the method described in \citet{fourier_feature_networks}.
			% We found this method gave better results than storing each differentiable texture as a raster matrix of pixels, as well as being faster to train.
			% During inference, you can store the results of this MLP as a raster image of arbitrarily large resolution, saving both time and VRAM.
			\label{fig:learnable_textures}
		\end{figure}

		%JUNE 12:
		%Defining what it is and how it's used.
		%
		%Justify why we're calling it neural neural texture. Unlike previous papers [citations], we learn a neural network instead of a grid of pixels. To be explicit we call it neural neural texture. %%%%%% Previous papers were more like latent textures but ours is different: it is a funciton.  Previous works were called "Neural Textures" but we're going to call ours "Neural Neural Textures" - a subtype of neural textures.

		% Using neural neural textures, we can encode features directly onto the surfaces of the 3d objects in a scene, which the image translation module uses to maintain visual consistency between states and viewpoints.
		Instead of feeding a scene $s$ directly into the image translator $\Tab$, we first apply learnable textures $\psi$ to its surfaces as shown in figure \ref{fig:main_diagram}.
		Previous works called these learnable textures ``neural textures'', and were parametrized by a discrete grid of differentiable texels. 
		% In contrast, we call our learnable textures ``neural nerual textures'', becuase our textures are a type of neural texture that is parametrized continuously over UV space using a neural fourier feature network.
		In contrast, we call our learnable textures ``neural nerual textures'', becuase our textures are parametrized continuously over UV space using a neural network.
		Like previous works, our neural textures $\psi\toL$ are learned jointly with our image translator $T$.

		These neural neural textures are represented implicitly by a neural network that maps UV values to RGB values. %, and are used to obtain projections $\pi$ from 3d scenes $s$.
		For each texture $\psi^i \in \psi\toL$, 
		\begin{equation}
			\psi^i: (u,v) \in\real^2 \rightarrow (r,g,b) \in\real^3 % \eqand 
		\end{equation}

		Given a scene $s$, we obtain projection $\pi$ by applying a texture $\psi^i \in \psi\toL$ to every pixel $(u,v,l) \in s$ individually, where texture index $i=I(l)$ is decided by the label value $l$ of that scene pixel.
		% Each pixel in the resulting projection $rgb\in\pi$ is the result of selecting a texture index $i$ based on the rank of label value $l$, then applying $\psi^i$ to the $u,v$ values.
		\begin{equation}
			% \pi=\psi\toL(s) \eqand rgb \in \pi = \psi^i(uv)
			% \pi_{n,c,x,y}=\psi^i(s_{n,1,x,y},s_{n,2,x,y}) \eqwhere 1\le n\le N,\ \  1\le c\le 3 ,\ \ 1\le x\le W,\ \ 1\le y\le H,\ \ \text{and} \ \ i=I(s_{n,3,x,y})
			% \pi_{n,c,x,y}=\psi^i(s_{n,1,x,y},s_{n,2,x,y})
			\pi_{[x,y]}=\psi^i(s_{[x,y,1]},s_{[x,y,2]})
		\end{equation}
		% where $1\le n\le N,\ \ 1\le c\le 3 ,\ \ 1\le x\le W,\ \ 1\le y\le H,\ \ \text{and} \ \ i=I(s_{n,3,x,y})$.
		% where $n\in \{1\dots N\},\ \ c\in \{1\dots 3\} ,\ x\in \{1\dots W\},\ y\in \{1\dots H\},\ \ \text{and} \ \ i=I(s_{n,3,x,y})$.
		where $ x\in \{1\dots W\},\ y\in \{1\dots H\}$, and texture index $i=I(l)$ where $l=s_{[x,y,3]}$ and subscript means multidimensional indexing.

		%DISCARDED TEXT:
		% A particular neural texture $\psi^i$ uses maps $uv \rightarrow rgb$. Similarly, the set of textures $\psi_{1\dots L}$ maps uvl values to rgb values: $\psi_{1\dots L}(uvl)=\psi_l(uv)$ where $uvl$ is the concatenation of a $uv$ vector with a label $l \in 1\dots L$.
		% TRITON's neural neural texture module lets it create features on the surfaces of objects, and thus take advantage of 3d geometric priors.

				% The neural texture applied to a particular scene pixel $uvl \in s$ is multiplexed by $l$, and the respective output pixel $rgb \in \pi = \psi_luv$. 
				% Texture projections $\pi$ are obatined by applying a set of $L$ different neural neural textures $\psi\toL$ to a UVL scene $s$ pixelwise.
				% [[IMPORTANT: THIS IS TECHNICALLY INCORRECT AS l IS A VALUE BETWEEN 0 and 1]]

				% where $i\in[1,L]\cap\mathbb{Z}$ is the rank of $l$
				% When applied to a UVL scene, it is applied to each $uvl \in s$ pixel individually where every pixel $rgb \in \pi = \psi$ 
				%What is it...then how do we use it...
				% (as the 2nd sentence)
				% Given a 3D scene $s$, we apply $\psi\toL$ to each UV to obtain RGB surface values...
				% project learnable RGB textures $\psi\toL$ to their surfaces to obtain projection $\pi$. 
				% 			Neural Neural Textures $\psi\toLuv$: Given an input scene $s$ depicting 3d objects with labels $1\dots L$, we apply a function on their surface's UV values with neural textures $\psi\toL$ to obtain projection $\pi$.

				% \begin{equation}
					% \pi=\psi\toL(s)
				% \end{equation}
				%

		%THIS WILL BE REPLACED BY SUMMARY SECTION
		% 	Given a batch of $N$ UVL Scenes $s_{1\dots B}$ with $L$ different label values, we project $L$ different neural textures $\psi_{1\dots L}$ onto that scene; obtaining projections $\pi_{1\dots B}$. This projection mechanism is illustrated in figure \ref{fig:uvl_explanation}.

		% Each neural texture $\psi^i$ is a function consisting of two components: a multi-layer perceptron $M^i$ and a static set of 256 random spatial frequencies $f_{1...256}$. 
		% Given a two-dimensional $uv$ vector, the spatial frequencies are used to generate a high dimensional embedding used as input for the MLP.
		% \begin{equation}
		% 		\psi^i(uv) = 
		% 		\
		% 		M^i\left(\left[
		% 			\sin \left( f_1     \cdot \left(u,v\right) \right) , \ 
		% 			\cos \left( f_1     \cdot \left(u,v\right) \right) , \ 
		% 			% \sin \left( f_2     \cdot \left(u,v\right) \right) , \ 
		% 			% \cos \left( f_2     \cdot \left(u,v\right) \right) , \ 
		% 			\dots \ 
		% 			\sin \left( f_{256} \cdot \left(u,v\right) \right) , \ 
		% 			\cos \left( f_{256} \cdot \left(u,v\right) \right)  \ 
		% 		\right]\right)
		% \end{equation}
		% where $uv$ and $f^i \in N_{ormal}(\mu=0,\ \sigma=10)$ are two dimensional vectors,
		%  $f^i uv$ represents their dot product, 
		% %  $rgb$ represents a 3d color vector,
		%  and $M^i$ maps a $uv$ embedding to an $rgb$ color value:
		%  $M^i: \real^{256} \rightarrow \real^3$.
		% Note that there is nothing special about 256 or 10: these hyperparameter values just happen to work nicely.

		We now discuss the implementation of our neural neural networks $\psi\toL$.
		Each neural neural texture $\psi^i$ is a function consisting of two components: a multi-layer perceptron $M^i$ and a static set of 256 random spatial frequencies $f_{1...256} \in \real ^ {256 \times 2}$. 
		Given a two-dimensional $(u,v)$ vector, the spatial frequencies are used to generate a high dimensional embedding of $(u,v)$ as input for $M^i$, where $M^i$ is a function $\real^{256} \to \real^{3}$ mapping that embedding to an $(r,g,b)$ color value:
		\begin{equation}
				\psi^i(u,v) = 
				\
				% M^i(
				% 	\sin( f_1     uv  ) , \ 
				% 	\cos( f_1     uv  ) , \ 
				% 	\sin( f_2     uv  ) , \ 
				% 	\cos( f_2     uv  ) , \ 
				% 	\dots \ 
				% 	\sin( f_{256} uv  ) , \ 
				% 	\cos( f_{256} uv  )  \ 
				% )
				M^i(
					\sin( g_1      ) , \ 
					\cos( g_1      ) , \ 
					\sin( g_2      ) , \ 
					\cos( g_2      ) , \ 
					\dots \ 
					\sin( g_{256}  ) , \ 
					\cos( g_{256}  )  \ 
				)
		\end{equation}
		where $g_k=f_k\cdot(u,v)$
		% , $f^i \in N_{ormal}(\mu=0,\ \sigma=10)$ is a two dimensional random gaussian vector,
		, $f_k \in \mathcal{N}(\mu=0,\ \sigma=10)$ is a two dimensional random gaussian vector,
		and $\cdot$ is the dot product operator.
		%  $rgb$ represents a 3d color vector,
		% :
		% $M^i: \real^{256} \rightarrow \real^3$.
		Note that there is nothing special about 256 or 10: these hyperparameter values just happen to work nicely.

		%Squeeze this down - we don't need explanations for why, that can go in the appendix.
		In practice, we found that TRITON learns faster and more stably with our neural neural textures than it does with raster neural textures. The generated textures are also smoother and less noisy. See the supplementary material for more details.
		% The trained fourier-feature textures also exhibit less noise than its raster counterparts. 
		% We think this is because they do not suffer as much. 
		% It cfrom aliasing artifacts in the UV maps, where the loss gradient might skip over pixels in texture-space when the UVL scene is zoomed out too far or raster resolution is too high. 
		% Case in point: if you have a raster texture with very high resolution, the loss gradient is less likely to be passed to a given texel because the chance that a given UV value in a scene will be rounded to that pixel's coordinates is very small.

		% UV Map projection
		% We use an RGB texture, parametrized by an MLP taking fourier features as an input.
		% We use a separate MLP for every texture.

	\subsection{Image Translation}
	\label{sec:image_translation}

	% 	\begin{figure}[H]
	% 		\begin{center}
	% 			\includegraphics[width=200pt]{../images/module_translator.pdf}
	% 		\end{center}
	% 		\caption{
	% 			A subset of figure \ref{fig:main_diagram} focusing on the image translation module $T$.
	% 		}
	% 		% For each texture, there is a separate Fourier feature network (Fourier features fed into a pixel-wise MLP) that will take in Fourier features as inputs, and output an RGB image using the method described in \citet{fourier_feature_networks}.
	% 		% We found this method gave better results than storing each differentiable texture as a raster matrix of pixels, as well as being faster to train.
	% 		% During inference, you can store the results of this MLP as a raster image of arbitrarily large resolution, saving both time and VRAM.
	% 		\label{fig:module_translator}
	% 	\end{figure}

		% TO PUT IN FORMULATION:
		% The simulation domain A (UVL maps concatenated to their texture projections) has 6 channels, whereas the real image domain B (photographs) has 3 channels. An element of A is $\ppis \in A$ and $p \in B$.
		% $p \in B$ is a batch of real photos,
		% and $\ppis\in A$ is a batch of simulated projection/uvl scene concatenations.

		% TODO: Why doesn't citep work? The references don't show the authors names...it used to work...did I mess up example.bib?

		% TODO: Get rid of the variable "C" and replace it with $\ppis$

		% TODO: Is l2 loss capitalized or lower case?

		%_i
		% Our image translation module $T$ is bidirectional: $\Tab$ translates images from domain A to domain B, and $\Tba$ translates images from domain B to domain A.
		Our image translation module $T$ is bidirectional: $\Tab$ translates sim to real (aka domain A to domain B), and $\Tba$ translates real to sim (aka domain B to domain A).
		\begin{equation}
			\ph=\Tab(\pis)  \eqand  \ppish=\Tba(p)
		\end{equation}

		$T$ is based on the image translation module used in \citet{surgical_video_translation}, which is a based on a variant \cite{surgical_image_translation} of MUNIT \cite{munit}. 
		$T$ uses the same network architectures as MUNIT, and also inherits its adversarial losses $\ell_{adv}$, cycle consistency losses $\ell_{cyc}$ and reconstruction losses $\ell_{rec}$.
		The main differences between our image translation model and MUNIT are that we use a different image similarity loss $\Omega$, and like \citet[]{surgical_image_translation} and \citet[]{surgical_video_translation} our style code is fixed (making it unimodal) and noise is injected into the latent codes during training to prevent overfitting. 
		During inference however, this intermediate noise is removed and our image translation module is completely determinstic. 
		% Apart from these differences, our image translation module has the same loss formulations and generator/encoder/discriminator network architectures as MUNIT.

		MUNIT translates images by encoding both domains $A$ and $B$ into a common latent space, then decoding them into their respective domains  $B$ and $A$.
		In our translation module, we define fake photos $\ph = \Tab(\pis) = G_B(E_A(\pis))$ and fake projection/uvl scenes $\ppish = \Tba(p) = G_A(E_B(p))$
			where $E_A, E_B, G_A, G_B$ are encoders and generators for domains A and B respectively.
		% We also define the outputs of these translators as fake photos $\hat{p} = T_{A\rightarrow B}(\pis)$ and fake projection/uvl scenes $(\hat{\pi},\hat{s}) = T_{A\rightarrow B}(\pis)$.

		% TODO: Choose one of the three equation representations. Which is best? The next paragraph will look less ugly when the equations are shorter.

		% For our purposes, MS-SSIM is a metric that measures the similarity of two BCHW tensors on a scale from 0 to 1.
		% Given two tensors $x$ and $y$, their similarity loss is $\mu\left((x-y)^2\right) - \text{msssim}(x,y)$. 
		We define an image similarity loss $\similar$ between two images $x,y$ that returns a score between 0 and 1, where 0 means perfect similarity:
		\begin{equation}
			\similar(x,y) = L_2(x,y)-\msssim(x,y) 
		\end{equation}
		This function combines mean pixel-wise $L_2$ distance (used in MUNIT) with multi-scale structural image similarity $\msssim$, introduced in \citep{msssim}. Note that this loss can also be applied to the latent representations obtained by $E_A$ and $E_B$, because like images those tensors are also three dimensional.

		% To ensure cycle consistency, we have an image similarity loss $\ell_{cyc}$ to ensure
		% % $c \approx G_A(E_B(G_B(E_A(\pis)))) = T_{B\rightarrow A}(T_{A\rightarrow B}(\pis)) = T_{B\rightarrow A}(\hat{p})$
		% $\ppis \approx T_{B\rightarrow A}(\hat{p})$
		% and vice versa
		% % $p \approx G_B(E_A(G_A(E_B(p)))) = T_{A\rightarrow B}(T_{B\rightarrow A}(\pis)) = T_{A\rightarrow B}(\hat{c}) = T_{A\rightarrow B}(\hat{\pi},\hat{s})$. 
		% $p \approx T_{A\rightarrow B}(\hat{\pi},\hat{s})$.
		% %I only want to use one of the two representations; either using T_AB or GA(EB) etc. Which one should I use?
		% We also impose an image similarity loss $\ell_{rec}$ to ensure that $\ppis \approx G_A(E_A\ppis)$ and $p \approx G_B(E_B(p))$, which is effectively an autoencoder loss. 
		% In addition, we also have content similarity loss $ \ell_{con}$ to make 
		% $E_A(\pis) \approx E_B(\ph)$ and $E_A(p) \approx E_A(\pish)$.
		% % $E_A(\pis) \approx E_B(\Tab(\pis))$ and $E_A(p) \approx E_A(\Tba(p))$.

		We have cycle consistency loss ${\ell_{cyc}  =  \similar\left(\ppis,T_{B\rightarrow A}\left(\hat{p}\right)\right)  +  \similar\left(p,T_{A\rightarrow B}\left(\hat{\pi},\hat{s}\right)\right)}$
		% $\ppis \approx T_{B\rightarrow A}(\hat{p})$
		% and vice versa
		% $p \approx T_{A\rightarrow B}(\hat{\pi},\hat{s})$.
		, similarity loss ${\ell_{rec} = \similar\left(\ppis,G_A\left(E_A\ppis\right)\right) + \similar\left(p,G_B\left(E_B\left(p\right)\right)\right)}$
		%  to ensure that $\ppis \approx G_A(E_A\ppis)$ and $p \approx G_B(E_B(p))$, 
			(which is effectively an autoencoder loss),   
		and content similarity loss ${ \ell_{con} = \similar\left(E_A\left(\pis\right), E_B\left(\ph\right) \right)    +   \similar\left(E_A\left(p\right),E_A\left(\pish\right)\right) }$.
		% $E_A(\pis) \approx E_B(\ph)$ and $E_A(p) \approx E_A(\pish)$.
		We also have adversarial losses $\ell_{adv}$ that come from two discriminators $D_A$ and $D_B$, targeting domains A and B respectively, using the LS-GAN loss introduced in \citet{lsgan}. 
		In total, our image translator loss is $\ell_T=\ell_{cyc}+\ell_{rec}+\ell_{con}+\ell_{adv}$.

		% PRETTIEST BUT TOO MUCH SPACE
		% 		Inheriting from MUNIT, we have three losses: cycle consistency loss $\ell_{cyc}$, content similarity loss $\ell_{con}$, and reconstruction loss $\ell_{rec}$ which functions as an autoencoder loss.
		% 		\begin{eqnarray}
		% \begin{center}
		% 	\begin{align}

		% 		\ell_{cyc}  =  \similar\left(\ppis,T_{B\rightarrow A}\left(\hat{p}\right)\right)  +  \similar\left(p,T_{A\rightarrow B}\left(\hat{\pi},\hat{s}\right)\right)

		% 		\ell_{con} = \similar\left(E_A\left(\pis\right), E_B\left(\ph\right) \right)    +   \similar\left(E_A\left(p\right),E_A\left(\pish\right)\right)

		% 		\ell_{rec} = \similar\left(\ppis,G_A\left(E_A\ppis\right)\right) + \similar\left(p,G_B\left(E_B\left(p\right)\right)\right)
		% 	\end{align}
		% \end{center}
		% 		\end{eqnarray}

		% In total, the loss for the image translator module $T$ is $\ell_T=\ell_{cyc}+\ell_{rec}+\ell_{con}+\ell_{adv}$.

		% % NOTE: THe other paper \citep{surgical_video_translation} didn't do into this much detail; instead putting most of these losses in teh appendix. I think we can cut this down? I'll check with michael first...right now I'll include everything.

	\subsection{Unprojection Consistency Loss}
	\label{sec:unprojection_consistency_loss}

		\begin{figure}[thbp]
		    \vspace{-10pt}
			\begin{center}
				\includegraphics[width=.8\textwidth]{../images/unprojection_consistency_loss_without_reprojection.pdf}
			\end{center}
			\vspace{-5pt}
			\caption{
				This is the unprojection consistency loss used in figure \ref{fig:main_diagram}.
				The Unprojection Mean and Reprojection nodes are not used in any losses, and help to illustrate the next section \ref{sec:texture_realism_loss}.
				% After calculating fake photos $\hat{p}_{1...B}$ from UVL scenes $s_{1...B}$ as seen in figure \ref{fig:main_diagram}, 
				% we can unproject fake photos $\hat{p}_{1...B}$ back into texture space to obtain $u_{1...B,\ 1...L}$ .
				% If the fake photos display consistent textures along their 3d surfaces, then the unprojected textures should be similar.
				% We can measure this similarity by taking the per-channel standard deviation of these unprojections, aka $\sigma(u)$, then averaging the result.
				% When the fake photos display similar surfaces, the unprojections will be similar making the unprojection consistency loss low.
				% Additionally, we can also calculate the average unprojection, which is useful for visualization purposes (it will look blurry if $\hat{p}$ is inconsistent).
			}
			\label{fig:unprojection_consistency_loss}
		\end{figure}

		To keep the object surfaces consistent between a batch of $N$ outputs, we impose a pixel-wise ``unprojection consistency loss'' between these surfaces by unprojecting them back into a common texture space.
		\begin{equation}
			u\toNL=\text{unproject}(\ph\toN,s\toN)   \eqand  \lUC \Rightarrow u_1 \approx u_2 \dots \approx u_N
		\end{equation}

		We sample a random batch of UVL scenes $S\toN$, and translate them into $\ph\toN$.
		Using geometric knowledge of the scene encoded in $S\toN$, we individually unproject each $\ph^i \in \ph\toN$ back into texture space, to obtain $N$ sets of $L$ unprojected textures $u\toNL$.
		We then use the per-pixel, batch-wise standard deviation of the unprojections as unprojection consistency loss $\lUC = \sigma(u\toNL)\toL$.
		This loss observes how different these unprojections are from one another, and thus how different the object surfaces appear between translations.

		Intuitively, if $\lUC$ were $0$, it would mean the object surfaces in translations $\ph$ appear exactly the same in every scene.

	%MOVE THIS TO APPENDIX
	% 	This loss is a fairly intuitive and effective way to maintain surface consistency. There are some caveats, howe/ver.
	% 	In order to produce a standard deviation greater than $0$, we must have a batch size greater than $1$.
	% 	Furthermore, environments that rarely show a particular region of a surface will rarely have $\lUC$ applied to that region because they're unlikely to show up twice in the same batch.
	% 	For this reason, $\lUC$ is most useful with large batch sizes.% because observing the same surface in multiple elements of a batch becomes more likely as the batch size increases. 
	% 	But because video memory limits our batch sizes, first-person environments might be more of a challenge for $\lUC$ - particularly when exploring multiple rooms (a task we haven't done though - should I mention this?)

		% To michael: Should we mention these caveats in this section, or mention them in the texture realism loss section? That section addresses all of these issues...

		% Also, you have to get lucky - two scene pictures have to show the same side of a cube for example. First person scenarios would be even worse.

		% %Unabbreviated
		% \begin{equation}
		% 	\unprojection_{c,U,V}^{n,i}=\expectation\ph_{c,x,y}^n 
		% 	\eqand \bar{\unprojection}^i_{c,u,v}   =   
		% 		\frac{1}{N} \lsum_{n=1}^N\unprojection^{n,i}_{c,u,v} 
		% 	\eqand 
		% 	\lUC=\frac{\lsum_{i=1}^L\lsum_{c=1}^3\lsum_{U=1}^{128}\lsum_{V=1}^{128}
		% 	\sqrt[]{\frac{1}{N}\lsum_{n=1}^N\left(\bar{\unprojection}^i_{c,u,v}-\unprojection^{i,n}_{c,U,V}\right)^2}}
		% 	{3\cdot128\cdot128\cdot L}
		% \end{equation}
		%
		%Abbreviated
		To formally define $\lUC$ we define unprojections $\unprojection^{1...N,1...L}$ and mean unprojections (aka recovered textures) $\bar{\unprojection}^{1...L}$ 
		with each $\unprojection^{n,i}\in\real^{3\times128\times128}$:
		\begin{equation}
			\unprojection_{U,V}^{n,i}=\expectation\ph_{x,y}^n 
			\eqand \bar{\unprojection}^i_{u,v}   =   
				\frac{1}{N} \lsum_n\unprojection^{n,i}_{u,v} 
			\eqand 
			%\lUC=\frac{\lsum_{i,U,V}
			%\sqrt[]{\frac{1}{N}\lsum_n\left(\bar{\unprojection}^i_{u,v}-\unprojection^{i,n}_{U,V}\right)^2}}
			%{128\cdot128\cdot L}
			\lUC=\mathbb{E}\left[
			\sqrt[]{\frac{1}{N}\lsum_n\left(\bar{\unprojection}^i_{u,v}-\unprojection^{i,n}_{U,V}\right)^2} \right]
		\end{equation}
		where $U=\lfloor128u\rfloor$ and $V=\lfloor128v\rfloor$ and $i=I(l)$ with
		%  $(u,v,l)=s^n_{1...3,x,y}$
		$u=s^n_{1,x,y}$, $v=s^n_{2,x,y}$, $l=s^n_{3,x,y}$
		for all 
		$n \in \{1...N\}$ , 
		$i \in \{1...L\}$ , 
		$x \in \{1...W\}$ , 
		$y \in \{1...H\}$ , 
		$U \in \{1...128\}$ , 
		$V \in \{1...128\}$.

	\subsection{Texture Realism Loss}
	\label{sec:texture_realism_loss}
        \vspace{-12pt}
		\begin{figure}[thbp]
			\begin{center}
				\includegraphics[width=250pt]{../images/texture_realism_ablation.pdf}
			\end{center}
			\vspace{-5pt}
			\caption{
				The neural texture looks more realistic with texture realism loss enabled. Note that the blocks show different letters because of different initializations and the symmetry of a cube - both solutions are valid texture assignments. 
			}
			\vspace{-10pt}
			\label{fig:texture_realism_ablation}
		\end{figure}

		To encourage the neural textures to look as realistic as possible, we try to make the projections look like the final output by introducing a ``texture realism'' loss $\lTR$. 
		$\lTR$ is an image similarity loss that makes $\pi$ look like its translation $\hat{p}$.
		% As it turns out, $\lTR$ is sufficiently powerful loss to generate surface-consistent translations even without $\lUC$, although the results are better when both losses are used.
		% This loss forces the neural textures to look realistic, instead of allowing them to have arbitrary colors.
		% In other words, $\lTR$ makes the neural textures more realistic by making their projections look more realistic, like the final outputs.
		\begin{equation}
			\lTR = \similar(\pi,\ph)
		\end{equation}
		% So far, we've taken an image translator $T$ and added neural textures $\psi$ and a surface consistency constraint $\lUC$. 
		% If we wanted to, could stop here - those formulations are sufficient to produce consistent sim-to-real translations.
		% However, we can improve this further.
		% Up until now we have not put any constraints on this neural texture. 
		Without $\lTR$, $\psi$ can look wildly different each time you train it.
		As seen in the top row of figure \ref{fig:texture_realism_ablation}, the neural texture looks very unrealistic – the colors are completely arbitrary.
		After adding $\lTR$ though, $\pih \approx \ph$, and $\psi$ to represents the albedo of the surfaces in the scene.
		% The image similarity loss used in $\lTR$ is calculated the same way image similarity losses are calculated in section \ref{sec:image_translation}: L2 loss minus MSSSIM.

		As it turns out, we can achieve surface-consistent translations just using $\lTR$, without $\lUC$.
		The the reverse is also true: $\lTR$ is sufficient to ensure surface consistency without $\lUC$.
		Using both, however, yields better results than either loss alone. 
		% For an explanation of why this is the case, see the appendix \ref{sec:texture_realism_theory}.

		With that said, there are some advantages to using $\lTR$ alone: when just using $\lTR$ you can use batch sizes as small as $N=1$, allowing you to save video memory. 
		Using $\lTR$ is also more numerically stable, as you no longer have to worry about whether the same surfaces are seen across multiple scenes in a batch as you would with $\lUC$.

		% We introduce a new loss, called texture realism loss that fixes this: $\lTR$ is an image similarity loss that encourages $\pi \approx \ph$.
		% This loss forces the neural texture to approximate the albedo of the surfaces in a scene.

		% These images could be regularized so that they're the same every time.
		% 
		% To do this we put a penalty on the latent texture that makes it $\pi$ look like its translation $\hat{p}$. 
		% 
		% We use l2 and msssim losses to do this.
		% 
		% This helps improve the translations.
		% 
		% There are some advantages over unprojection consistency though: 
		% smaller batch sizes (even a batch size of 1, which UC cant do),
		% you don't have to get lucky (the more chaotic, or even the more zoomed in a dataset becomes the less likely there is to be any collisions)

	%===============================================================================

\section{Experiments}\label{sec:results}

\subsection{Datasets}
Here we detail the datasets used in our experiments. 
Every dataset has two sets of images: UVL scenes and real photographs. They are all unpaired, and have RGB values ranging from 0 to 1.

The first dataset to mention is the three-block dataset.  
It has 300 photos and 2000 UVL scenes. It features a table with three alphabet blocks on a table, which are randomly translated along the X and Y axes and rotated randomly along the Z axis. Each photograph features these three cubes with a random position and rotation.

[[Insert photographs and UVL maps here; a digram showing each dataset we describe here and 1 or two samples of each domain in each dataset]]

% In this paper, we'll mostly focus on a hand-collected dataset that involves three alphabet blocks that are moved across a table.
% Alphabet blocks were chosen because they all have the same geometry (a cube), but have very distinct semantic textures that will make temporal inconsistency obvious to the human eye.
% This dataset is analogous to data that would be collected for a third-person robotic grasping task.
% However, other objects (such as cans of soda, apples, cloves of garlic, etc) are also for generating other results in this paper. ((TODO: Detail the five object tests and five cube tests and other stuff like shiny cans etc)).

% There are two components to this dataset: Simulated data in the form of UVL scenes, and photographs.
% The camera in the simulation is positioned in approximately the same place as the real camera, with approximately the same field of view.
% In both domains of this dataset, the camera never moves. Instead, the objects are randomly positioned and rotated on the table.
% In both the simulation and in real life, these objects are randomly translated along the x and y-axis of the table, as well as rotated on the z-axis.

% There are 200 photographs in this dataset, and they are recorded as RGB jpg images.
% There are 3000 synthetic images in this dataset, in the form of UVL images.
% These UVL images are also encoded as RGB images, but provide information about the geometry of a scene.
% UVL images are composed of both a UV map and a label map and are summarized in figure \ref{fig:uvl_explanation}.
% To avoid positional rounding errors, UVL images are encoded as EXR files: an image format that supports floating-point values.

% The pixels in a UVL image are visualizable in RGB space because, like RGB images, each pixel is a vector with 3 values. The first two channels, U and V, are floating point values between 0 and 1. The L channel is also a floating point value between 0 and 1, it has a second meaning. There are only a finite number of label values (corresponding to the number of learnable textures). The rank of a particular L value is used to index these textures.

% In the three-cube dataset we have 4 textures, and UVL scenes contain label values $l \in [\frac{0}{3}, \frac{1}{3},\frac{2}{3}, \frac{3}{3}]$. A pixel in a UVL scene $l=\frac{2}{3}$ will get indexed to the third texture, because $\frac{2}{3}$ is the third smallest label value.

% %===============================================================================

% Domain A and B
% p hat
% s hat

\subsection{Qualitative Evaluation Results}
\begin{figure}[thbp]
    \vspace{-30pt}
	\begin{center}
		\includegraphics[width=\textwidth]{../images/frame_inconsistency_diagram.pdf}
	\end{center}
	% Each outline color corresponds to the same cube.
	% The same cubes are marked with the same colored outlines across all photos.
	% For the same cube, we expect to see the same number on the top of the cube regardless of its placement.
	\caption{
		Each row of images shows a different random placement of objects in the scene.
		Each column of images is a different image translation algorithm.
		The same cube is marked with the same color outline across all photos.
		TRITON is consistent: for example, the cube outlined as yellow always has an 8 on the top.
		Meanwhile, CycleGan for example shows a 4 on the top of that cube in the top two rows, but an 8 and a 7 in the next two rows.
		Because of the symmetry of the cubes, and the unsupervised training, it is arbitrary which synthetic cube gets which symbol after training - but whatever it chooses, it should be consistent between placements.
		CUT and CycleGan are not consistent, while TRITON is.
		% TODO: Update the images in this to have the latest training iteration...
		}
		\label{fig:frame_inconsistency_diagram}
	\end{figure}

    \begin{figure}[thbp]
    	\begin{center}
    		\includegraphics[width=\textwidth]{../images/blurry_recovery.pdf}
    	\end{center}
    	\caption{
    		The recovered texture is more crisp when we have surface consistency losses $\lUC$ and $\lTR$ enabled.
    		[[TODO: include cyclegan and other algorithms here]]
    	}
    	\label{fig:blurry_recovery}
    \end{figure}
	% 	Alphabet Three							
	% 	Masked				Unmasked			
	% 	LPIPS		L2		LPIPS		L2	
	% 	Sync	Unsync	Sync	Unsync	Sync	Unsync	Sync	Unsync
	% Pure MUNIT	0.0437	0.0373	0.00500	0.00490	0.434	0.402	0.0425	0.0392
	% Texture Only	0.0442	0.0386	0.00586	0.00589	0.285	0.255	0.0364	0.0310
	% Texture Reality	0.0286	0.0267	0.00479	0.00478	0.117	0.1157	0.0123	0.0122
	% CycleGAN			TODO: Sync with respect to L2, so sync is always ≥ unsync					
	% CUT				TODO: Google Synthetic Dataset

	\subsection{Ground Truth Comaprison}

	%CREATED WITH https://www.tablesgenerator.com/# -- this website is wonderful!!
\begin{table}[H]
    \caption{Ground truth comparison evaluations for TRITON. Masked and Unmasked mean xxxxxx. Sync and Unsync mean xxxxx. L2 is xxxx LPIPS is xxxxxx. TRITON is xxxxxxx. }
	\resizebox{\textwidth}{!}{
        \begin{tabular}{lllllllll}
        \toprule
        & \multicolumn{4}{c}{Masked} & \multicolumn{4}{c}{Unmasked}  \\
        \cmidrule(lr){2-5} \cmidrule(lr){6-9} & \multicolumn{2}{c}{LPIPS $\downarrow$} & \multicolumn{2}{c}{L2 $\downarrow$}  & \multicolumn{2}{c}{LPIPS $\downarrow$}  & \multicolumn{2}{c}{L2 $\downarrow$}   \\
        \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
        Algorithm  & Sync    & Unsync   & Sync      & Unsync    & Sync   & Unsync  & Sync     & Unsync \\
        \midrule
        TRITON w/o $\psi, \lUC, \lTR$ & 0.0437                        & 0.0373                        & 0.00500                        & 0.00490                        & 0.434                        & 0.402                        & 0.0425                        & 0.0392                        \\
        TRITON w/o $\lUC, \lTR$       & 0.0442                        & 0.0386                        & 0.00586                        & 0.00589                        & 0.285                        & 0.255                        & 0.0364                        & 0.0310                        \\
        \textbf{TRITON}                        & \textbf{0.0286}               & \textbf{0.0267}               & \textbf{0.00479}               & \textbf{0.00478}               & \textbf{0.117}               & \textbf{0.1157}              & \textbf{0.0123}               & \textbf{0.0122}               \\
        \midrule
        CycleGAN                      & {\color[HTML]{000000} 0.0450} & {\color[HTML]{000000} 0.0379} & {\color[HTML]{000000} 0.00559} & {\color[HTML]{000000} 0.00554} & {\color[HTML]{000000} 0.602} & {\color[HTML]{000000} 0.593} & {\color[HTML]{000000} 0.0425} & {\color[HTML]{000000} 0.0423} \\
        CUT                           & 0.0469                        & 0.0382                        & 0.00553                        & 0.00526                        & 0.540                        & 0.534                        & 0.0637                        & 0.0629\\
        \bottomrule
        \end{tabular}
	}
\end{table}

[[Figures: Show these results and explain why we have a mask on some of them]]

TODO: Explain what sync and unsync mean, retrieve the results for CycleGAN and CUT

\subsection{Sim2real Transfer for Robot Learning}
\begin{figure}[thbp]
    \centering
    \vspace{-10pt}
    \includegraphics[width=\textwidth]{images/sim2real_framework.pdf}
    \vspace{-20pt}
    \caption{Sim2real framework of using TRITON. We evaluate in two settings: Seen and Unseen. In Unseen, we use a new camera viewpoint to train and evaluate the policy. This new viewpoint is not used for training TRITON. }
    \vspace{-5pt}
    \label{fig:sim2realframework}
\end{figure}
In this section we demonstrate how TRITON improves sim2real transfer for robot learning. To do this, we first train TRITON on a dataset consist of unpaired UVL scenes from simulator and photos from the real robot. After that, we train the robot policy in the simulator only. The input of the policy is the fake photo $\hat{p}$ translated from the simulated UVL scene using TRITON's translator $T_{A\to B}$. Finally, the trained policy is evaluated on the real robot, taking real photos $p$ as input. 
Ideally, a good image translation model leads to less gap between $p$ and $\hat{p}$, so that the policy trained on synthesized photos will transfer to real domain seamlessly.

\paragraph{Task} We use a robot pose imitation task for our experiments. Given the input of a photo of real robot pose, the policy outputs robot controls to replicate that target pose. We measure the angular error between the replicated and target poses as our evaluation metric:$\sqrt{\sum_{j} (\hat{a}_j - a_j)^2}$,
where $\hat{a}_j$ and $a_j$ are replicated and target joint angle respectively, and $j$ is the joint index. We use an xArm robot which has 7 joints. Moreover, we attach patterns and tapes on it to make its texture complex.

\paragraph{Robot policy and Baselines}
Since sim2real provides abundant training data using the simulator, we use behavior cloning which is enough to train a good robot policy.
The policy is implemented by CNN~\cite{dqn}.
We compare TRITON against two baseline image translation methods: CycleGAN and CUT, by replacing TRITON with each baseline method respectively in the above pipeline. The robot policy is the same for all methods.

\paragraph{Evaluation Settings} We introduce two main evaluation settings, \textbf{Seen} and \textbf{Unseen}. In \textbf{Seen} setting, the policy is trained and tested using the same camera viewpoint that is used in training image translator. Whereas in \textbf{Unseen} setting, we introduce a camera at a new viewpoint for training and testing the policy. That is, the image translator has to generate the fake photos out of its training distribution (seen camera viewpoints), which becomes a challenging for the translator. We also vary input image resolution to show the power of photorealistic images in higher resolutions.

\begin{figure}[tbhp]
    \centering
    \vspace{-8pt}
    \includegraphics[width=\textwidth]{images/sim2real_res.pdf}
    \vspace{-20pt}
    \caption{Evaluation results of sim2real robot learning. We compare TRITON against two baselines and one TRITON variant w/o $\ell_{UC}$ across 3 different input image resolutions. TRITON outperforms CycleGAN~\cite{} and CUT~\cite{} consistently across (a) all resolutions and (b) Unseen viewpoints. Results in (a) are from Seen setting. Results in (b) are from 336x336 resolution. }
    \label{fig:sim2real}
    \vspace{-8pt}
\end{figure}

\paragraph{Results}
In Figure~\ref{fig:sim2real} we show the evaluation results of sim2real transfer. Both evaluations show that TRITON outperforms CycleGAN and CUT consistently. In Figure~\ref{fig:sim2real}(a), we find that TRITON continuously improves from low to high resolutions due to the ability to generate more photorealistic images compared to the baselines. This is because in low resolution (eg. $84\times 84$) the details in images are not important, whereas in high resolution ($336\times 336$), artifacts from imperfect translator are clearer and do harm to policy learning.
From Figure~\ref{fig:sim2real}(b), TRITON outperforms the others in Unseen setting, showing the power of xxxxxx.
In the comparison against TRITON w/o $\ell_{UC}$, TRITON outperforms this variant consistently except the lowest resolution. This again shows the effectiveness of $\ell_{UC}$ to generate photorealistic images, and such high quality images are important in higher resolutions for sim2real transfer.

%===============================================================================

\section{Limitaions}
\label{sec:Limitations}

Can't handle transparent objects.

A limited number of objects can be learned at once before supervision is necessary with color loss (but there are workarounds).

Takes a long time to train.

%===============================================================================

% \section{Citations} 
% \label{sec:citations}

% Citations can be made using either \textbackslash citep\{\} or \textbackslash citet\{\}, depending on the appropriateness. To avoid the citation moving to the next line, it is often a good practice to replace the space before with a tilde (\~{}) character.
% Example 1: ``CoRL is the best conference ever~\citep{fourier_feature_networks}.''
% Example 2: ``\citet{fourier_feature_networks} proved, both theoretically and numerically, that CoRL is the best conference ever.''

%===============================================================================

\clearpage
% The acknowledgments are automatically included only in the final and preprint versions of the paper.
\acknowledgments{}

%===============================================================================

\bibliography{example}  % .bib

\clearpage

\section{Appendix} 
\label{sec:appendix}

    \subsection{Unsorted}
     	These neural textures are not stored in raster form during training. Instead, as illustrated in figure \ref{fig:learnable_textures}, they're paramaterized by pixel-wise fourier feature networks \cite{fourier_feature_networks} that take in UV coordinates and returns RGB colors. During inference, however, these textures are rasterized to save both time and video memory.

	\subsection*{Unprojection Recovery Resolution Effect}

		Also, the resolution of the unprojection matters:

		\begin{figure}[H]
			\begin{center}
				\includegraphics[width=400pt]{../images/unprojection_resolution_comparison.pdf}
			\end{center}
			\caption{
				The resolution of an unprojection matters for unprojection consistency loss. The larger it is, the more precise the alignment will be but the less likely a given UV value is to be assigned a loss greater than 0, lowering numerical instability.
			}
			\label{fig:unprojection_resolution_comparison}
		\end{figure}

	\subsection*{Texture Realism Theory}
		\label{sec:texture_realism_theory}

		THIS SECTION IS UNPOLISHED, BUT CONTAINS  THE NESSECARY CONTENT.
		I'll delete this section if you don't think its important enough to keep.

		Texture realism loss equation:
		$L_{TR}=L_{TR_{l2}}+L_{TR_{msssim}}$
		where 
		$L_{TR_{msssim}} = -msssim(\pi,\hat{p})$ and
		$L_{TR_{l2}} = \frac{
			\lsum_{i=1}^B {
			%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
			\left( \pi-\hat{p}  \right)^2
				} 
		}
		{B}$

		There's a good reason for this: it's almost equivalent to unprojection consistency.
		In fact, it can even be used on its own without unprojection consistency loss (both are sufficient to translate things.)

		(Give theoretical explation)
		$
		\ell_{UC} = 
		\
		\
		\mu\left(\sigma\left(u_{1\dots B,\ 1\dots L}\right)_{1 \dots L}\right) = 
		\sqrt{
		\frac{
			\lsum_{i=1}^B {
				%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
				\left( \hat{p}-\bar{p}  \right)^2
				} 
			}
			{B}
		}
		$
		Using $\bar{p}$ from figure \ref{fig:unprojection_consistency_loss}

		Becuase $\bar{p}$ is created from the average recovered texturs of $\hat{p}$, $\bar{p} \approx \hat{p}$.

		Becuse of texture realism loss, $\pi \approx \hat{p}$.

		Therefore, $\bar{p} \approx \hat{p} \approx \pi$

		% The l2 component of the texture realism loss equation:
		% $ \
		% \
		% \
		% \
		% \ \ell_{TR_{l2}}=\sum\limits_{i=1}^B {
		% 	%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
		% 	\left( \pi-\hat{p}  \right)^2
		% 	 } 
		% $

		Substituting one projected texture for another, aka $\pi$ for $\bar{p}$, we get 
		$ \
		\
		\
		\
		\ \ell_{TR_{l2}}
		\approx
		\frac{\sum\limits_{i=1}^B {
			%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
			\left( \hat{p}-\pi  \right)^2
			}}{B} = \ell_{UC}^2
		$

		\subsection{Color Loss}

		TRITON is very robust to geomtric imperfections. However, sometimes it takes this a bit too far - it can overfit textures onto the wrong objects and still maintain consistency.
		In one example, it manages to map an apple texture onto a cube.

		To help TRITON correctly match textures to objects, we introduce a small amount of supervision by adding a color loss. I'm going to stop writing this because I don't know if I even want to keep this in the paper...you can achieve the same result by running the algorithm multiple times, and so its not strictly nessecary.

		ASK MICHAEL: I'd rather not mention this entire section and run out the page limit...it's not very important, and not used in the robot experiments. It's only used in the experiment with the five items including the apple. Can we add it to the appendix instead?

\section*{Graveyard}

Here I put things I won't include in the paper or appendix, but want to keep for reference until submission

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=200pt]{../images/dataset_explanation.pdf}
	\end{center}
	\caption{
		Compatible datasets have two sets of images: a set of UVL scenes and a set of photographs.
		TRITON is a sim-to-real image translation algorithm, taking in synthetic UVL scenes and aiming to create realistic fake photographs. 
		Being an unsupervised algorithm, there are no correspondences between synthetic and real images.
		In this dataset, three alphabet cubes are moved around the table. 
		The synthetic and rear cameras have the same position and field of view.
		There are many more UVL scenes than photos because synthetic scenes can be generated automatically whereas photos require real-world labor.
		}
	\label{fig:dataset_explanation}
\end{figure}

\section*{Things to Ask Michael}

QUESTIONS FOR MICHAEL:
\begin{itemize}
	\item - TODO: WE MAINTAIN WHAT WE NOW CALL "SURFACE CONSISTENCY" instead of view consistency! Or texture consistency? Etc? What should we call it?
\end{itemize}

\end{document}